### 2023-02-20 outputs/final_paper.qmd
Explored GDP data

---
title: "TBD"
title-block-banner: true
#subtitle: "TBD"
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3014-final-paper.git"
author: "Christina Wei"
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up ####

library(tidyverse) # a collection of data-related packages
library(knitr) # for kable
library(kableExtra)
library(here)

#### Read in data ####

cleaned_world_economic_outlook_data = read_csv(
  file = here("inputs/data/cleaned_world_economic_outlook_data.csv"),
  show_col_types = FALSE
)
reference_country = read_csv(
  file = here("inputs/data/reference_country.csv"),
  show_col_types = FALSE
)

filtered_countries = c("USA", "CAN")
```

# Hypotheses

The tale of two countries - USA and China's influence in the world

The bigger you are measured by GDP, the more you control the world

Currently the world is driven by two countries

-   Inflations

-   Pegged currency?

-   Pie chart

Proving the digital divide: the richer is getting richer, the poor is getting poorer applies at country level

Divide up between rich and poor (top 5 and bottom 5?)

Poverty vs. GDP - any correlations?

Income inequality??

Unemployment

```{r}
top_5_countries_2020 =
  cleaned_world_economic_outlook_data |>
    filter(year == 2020) |>
    arrange(desc(GDP)) |>
    slice(1:5) |>
    select(country_iso) |>
    distinct()

top_5_countries_2020 =
  top_5_countries_2020$country_iso

bottom_5_countries_2020 =
  cleaned_world_economic_outlook_data |>
    filter(year == 2020) |>
    arrange(GDP) |>
    slice(1:5) |>
    select(country_iso) |>
    distinct()

bottom_5_countries_2020 =
  bottom_5_countries_2020$country_iso
```

```{r}
#| message: false
#| echo: false

cleaned_world_economic_outlook_data |>
  filter(country_iso %in% top_5_countries_2020) |>
  ggplot(aes(x = year, y = GDP, colour = country_iso)) +
  geom_line()

```

```{r}
#| message: false
#| echo: false

## Create a visual for 2020 GDP to show size differences between countries

# Create a 2020 GDP table, showing the top 20 countries + other countries summed up
gdp_data_2020 = 
  cleaned_world_economic_outlook_data |> 
  filter(year == 2020) |>
  select(country_iso, year, GDP)

gdp_data_2020 =
  merge(gdp_data_2020, reference_country, by = "country_iso") |>
  select(-country_iso) |>
  mutate(
    country = 
      recode(
        country,
        "T¸rkiye" = "Turkiye",
        "Islamic Republic of Iran" = "Iran",
      )
  ) |>
  arrange(desc(GDP))

gdp_2020_other = sum(slice(gdp_data_2020, 21:n())$GDP, na.rm = TRUE)

gdp_data_2020 =
  gdp_data_2020 |>
  slice(1:20) |>
  arrange(GDP) |>
  add_row(country = "Other", year = 2020, GDP = gdp_2020_other)

# Add percentage and cumulative percentage
# Help from https://r-charts.com/part-whole/pie-chart-percentages-ggplot2/
gdp_data_2020 = 
  gdp_data_2020 |>
  group_by(year) |>
  mutate(gdp_perc = GDP / sum(GDP, na.rm = TRUE)) |>
  mutate(gdp_cumperc = (cumsum(GDP) / sum(GDP, na.rm = TRUE)))

# Set order for graph
gdp_data_2020$country =
  factor(
    gdp_data_2020$country,
    levels = c("Other",gdp_data_2020$country[1:length(gdp_data_2020$country)-1])
  )

```

```{r}
#| message: false
#| echo: false

# Create bar chart based on GDP size
gdp_data_2020 |>
  ggplot(aes(country, GDP)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(
    name = "GDP Current Price in US Dollars (billions$)",
    limits = c(0, 21500)
  ) + 
  theme_minimal() +
  geom_text(
    aes(label = scales::percent(gdp_perc, accuracy = 0.1)),
    hjust = -0.1,
    size = 2
  ) + 
  labs(
    x = "Country",
    title = "Top 20 Countries with Highest GDPs in 2020"
  )

```