### 2023-03-20 outputs/paper/paper.qmd
update final project to investigate my own behavioural patterns in PhD

---
title: "TBD"
title-block-banner: true
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/TBD.git"
author: Christina Wei
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up and read in data ####

# Load packages 
library(tidyverse)
library(here)

# read in cleaned data

cleaned_time_sheet = read_csv(
  file = here("outputs/data/cleaned_time_sheet.csv"),
  show_col_types = FALSE
)
```

# Introduction

# Data

-   Timesheet tracked in Toggl Track

    -   Explain variables, categories

-   Sleep data - automatically tracked through Fitbit

-   Deliverable data - created at the beginning of each term to track major deliverables

-   Data cleaning procedures

-   Sources of error - retrospective entering of data, or forgetting to track; categorization arbitrary and may not reflect the actual work

# Results

**Timesheet**

-   From 2022-08-15 to 2023-03-13

-   582 entries

```{r}
min(cleaned_time_sheet$start_date)
max(cleaned_time_sheet$end_date)
nrow(cleaned_time_sheet)
```

```{r}

# Piecharts: https://r-charts.com/part-whole/pie-chart-ggplot2/
# Update this to %
cleaned_time_sheet |>
  group_by(type) |>
  summarize(effort_hours = as.numeric(sum(duration)) / 3600) |>
  ggplot(aes(x = "", y = effort_hours, fill = type)) +
  geom_col(color = "black") +
  coord_polar(theta = "y")
```

**Timesheet**

-   How long does it take to work on a paper?

-   When am I most productive?

-   What is the regular schedule?

```{=html}
<!-- -->
```
-   Study pattern?

```{=html}
<!-- -->
```
-   Monday to Sunday

-   Timing of the day

-   Length for type of work

-   How many things are happening concurrently?

Sleep score & time sheets

Actual vs. estimated effort

Behavioural patterns?

Model:

-   input - type of assignment, expected hours, done to date

-   process

-   output - recommended study schedule?

# Discussion

-   Missing - social aspects, quality of work

# Conclusion