### 2023-03-27 outputs/paper/paper.qmd
updated scripts & visuals for daily routine

---
title: "TBD"
title-block-banner: true
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/TBD.git"
author: Christina Wei
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

```{r}
#| message: false
#| echo: false
#| warning: false

#### Workspace set-up and read in data ####

# Load packages 
library(tidyverse)
library(here)
source(here("scripts/03-helper_functions.R"))

# read in cleaned data

cleaned_time_sheet = read_csv(
  file = here("outputs/data/cleaned_time_sheet.csv"),
  show_col_types = FALSE
)
```

# Introduction

# Data

-   Timesheet tracked in Toggl Track

    -   Explain variables, categories

-   Sleep data - automatically tracked through Fitbit

-   Deliverable data - created at the beginning of each term to track major deliverables

-   Data cleaning procedures

-   Sources of error - retrospective entering of data, or forgetting to track; categorization arbitrary and may not reflect the actual work

# Results

## Timesheet Data

### Data Charateristics

**Timesheet**

-   From 2022-08-15 to 2023-03-13

-   582 entries

```{r}
#| message: false
#| echo: false
#| warning: false

min(cleaned_time_sheet$start_date)
max(cleaned_time_sheet$end_date)
nrow(cleaned_time_sheet)
```

```{r}
#| message: false
#| echo: false
#| warning: false

cleaned_time_sheet |>
  group_by(type) |>
  summarize(effort = sum(effort_hours)) |>
  arrange(desc(effort))

# Piecharts: https://r-charts.com/part-whole/pie-chart-ggplot2/
# Update this to %
cleaned_time_sheet |>
  group_by(type) |>
  summarize(effort = sum(effort_hours)) |>
  mutate(effort_pct = effort / sum(effort)) |>
  ggplot(aes(x = "", y = effort_pct, fill = type)) +
  geom_col(color = "black") +
  coord_polar(theta = "y")
```

### Trend Over Time

```{r}
#| message: false
#| echo: false
#| warning: false

# Weekly trend
# https://stackoverflow.com/questions/67847174/convert-daily-data-into-weekly-data-and-summarize-multiple-columns-in-r

#line graph overall
generate_time_series_graph(
  ggplot_data = 
    cleaned_time_sheet |>
    group_by(wk_index = floor_date(start_date, "week")) |>
    summarize(effort = sum(effort_hours)) |>
    ggplot(aes(x = wk_index, y = effort)),
  xlabel = "Time",
  ylabel = "Weekly Effort (in hours)",
) +
geom_smooth(method = lm, se = FALSE)

# bar graph comparison over time
generate_compare_bar_graph(
  ggplot_data =
    cleaned_time_sheet |>
    filter(type %in% c("TA", "Research", "Course Work")) |>
    group_by(wk_index = floor_date(start_date, "week"), type) |>
    summarize(effort = sum(effort_hours)) |>
    ggplot(aes(x = wk_index, y = effort, fill = type)),
  position = "stack", 
  xlabel = "Week",
  ylabel = "Effort (in hours)"
)
```

### Daily Patterns

```{r}
#| message: false
#| echo: false
#| warning: false

# Duration of each session
cleaned_time_sheet |>
  group_by(start_date) |>
  summarize(
    effort = mean(effort_hours),
    sessions = n()
  ) |>
  ungroup() |>
  ggplot(aes(x = "", y = effort)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(width = 0.1) +
  labs(
    x = "",
    y = "Duration (in hours)"
  ) +
  theme_minimal()
```

```{r}
#| message: false
#| echo: false
#| warning: false

# Hours of each day
hourly_calc <- 
  cleaned_time_sheet |>
  slice(rep(1:n(), each = 24)) |>
  mutate(hour_value = rep(0:23, times = nrow(cleaned_time_sheet))) |>
  mutate(
    worked = (start_date == end_date & 
                hour(start_time) <= hour_value & 
                convert_hms_to_hours(end_time) > hour_value
              ) |
              (start_date < end_date & 
                 (convert_hms_to_hours(end_time) > hour_value |
                    hour(start_time) <= hour_value)
               )
    ) |>
  filter(worked) |>
  select(-worked)
```

```{r}
#| message: false
#| echo: false
#| warning: false

generate_compare_bar_graph(
  ggplot_data =
    hourly_calc |>
    filter(type %in% c("Course Work", "Learning", "Research", "TA")) |>
    group_by(hour_value, type) |>
    count() |>
    ungroup() |>
    mutate(pct = n / sum(n)) |>
    ggplot(aes(x = hour_value, y = pct, fill = type)),
  xlabel = "Hour of the Day",
  ylabel = "Percentage of Occurrence",
  position = "stack",
  yscale = scales::percent
)

```

### Weekly Patterns

```{r}
#| message: false
#| echo: false
#| warning: false

daily_data <- 
  cleaned_time_sheet |>
  mutate(wkday = wday(cleaned_time_sheet$start_date, label = TRUE)) |>
  group_by(start_date, wkday, type) |>
  summarize(effort = sum(effort_hours)) |>
  ungroup()
```

```{r}
#| message: false
#| echo: false
#| warning: false

daily_data |>
  group_by(start_date, wkday) |>
  summarize(effort = sum(effort)) |>
  ggplot(aes(x = wkday, y = effort)) +
  geom_boxplot() +
  stat_summary(fun.y="mean",color="red") +
  geom_jitter(width = 0.15) +
  labs(
    x = "Day of the Week",
    y = "Effort (in hours)"
  ) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 13, by = 1))

daily_data |>
  group_by(start_date, wkday) |>
  summarize(effort = sum(effort)) |>
  group_by(wkday) |>
  summarize(mean_effort = mean(effort), med_effort = median(effort))
```

### Concurrent Work

```{r}
#| message: false
#| echo: false
#| warning: false

cleaned_time_sheet |>
  group_by(start_date, project) |>
  summarize(projects = n()) |>
  group_by(start_date) |>
  summarize(projects = n()) |> 
  ggplot(aes(x = "", y = projects)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(width = 0.1) +
  labs(
    x = "",
    y = "Number of Projects Worked on Per Day"
  ) +
  theme_minimal()

cleaned_time_sheet |>
  group_by(start_date, project) |>
  summarize(projects = n()) |>
  group_by(start_date) |>
  summarize(projects = n()) |> 
  group_by(projects) |>
  count() |>
  ungroup() |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = projects, y = pct)) +
  geom_bar(stat = "identity") +
    labs(
    x = "Number of Projects Worked on Per Day",
    y = "Percentage of occurrences"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent)
```

## **Research**

### Overall

```{r}
#| message: false
#| echo: false
#| warning: false

cleaned_time_sheet |>
  filter(type == "Research") |>
  mutate(project = ifelse(project %in% c("Reading", NA), "Research-General", project)) |>
  group_by(project) |>
  summarize(effort = sum(effort_hours)) |>
  mutate(pct = effort / sum(effort)) |>
  arrange(desc(effort))

cleaned_time_sheet |>
  filter(type == "Research") |>
  group_by(start_date) |>
  summarize(
    effort = mean(effort_hours),
    sessions = n()
  ) |>
  ungroup() |>
  ggplot(aes(x = "", y = effort)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(width = 0.1) +
  labs(
    x = "",
    y = "Duration (in hours)"
  ) +
  theme_minimal()

generate_compare_bar_graph(
  ggplot_data =
    hourly_calc |>
    filter(type %in% c("Research")) |>
    group_by(hour_value, type) |>
    count() |>
    ungroup() |>
    mutate(pct = n / sum(n)) |>
    ggplot(aes(x = hour_value, y = pct, fill = type)),
  xlabel = "Hour of the Day",
  ylabel = "Percentage of Occurrence",
  position = "stack",
  yscale = scales::percent
)
```

### **Literature Review Paper**

-   How long does it take to work on a full paper?

```{r}
#| message: false
#| echo: false
#| warning: false

generate_time_series_graph(
    ggplot_data = 
      cleaned_time_sheet |>
      filter(project == "Lit Review") |>
      group_by(wk_index = floor_date(start_date, "week")) |>
      summarize(effort = sum(effort_hours)) |>
      ggplot(aes(x = wk_index, y = effort)),
    xlabel = "Week",
    ylabel = "Effort (in hours)",
  )

cleaned_time_sheet |>
  filter(project == "Lit Review") |>
  group_by(start_date) |>
  summarize(
    effort = mean(effort_hours),
    sessions = n()
  ) |>
  ungroup() |>
  ggplot(aes(x = "", y = effort)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(width = 0.1) +
  labs(
    x = "",
    y = "Duration (in hours)"
  ) +
  theme_minimal()
```

## **Type of Work**

### Technical Work

### Research

## **Actual vs. Estimated Effort**

## Parking Lot

-   When am I most productive?

-   What is the regular schedule?

-   Study pattern?

-   Monday to Sunday

-   Length for type of work

-   Difference in work patterns between writing & technical work

Actual vs. estimated effort

Behavioural patterns?

Work life balance?

# Model

Model1:

-   input - timesheet for the day + sleep score

-   process

-   output - effectiveness score: break down to different aspects

Model2: Extrapolate actual from expected hours?

# Discussion

-   Missing - social aspects, quality of work

-   Weekends / timeoff

-   Different types of flow - technical vs. research

-   Find the focus each day - try to stay with 1 or 2 topics only

-   Prefer to work on morning / early afternoon, and after dinner

# Conclusion

# Appendix {.unnumbered}

# Reference